Approximation as Differentiation: Notes

### 2025-06-06 notes/language.tex
C -> Sem for semantic category. Fix up Makefile for MacOS.

To give the semantics for the language defined in \figrefTwo{syntax}{typing}, we fix a bicartesian closed
category $\cat{C}$ with finite products $(\times, \One)$, finite coproducts $(\coprod, \Zero)$, exponentials
$\internalHom{X}{Y}$, evaluation morphisms $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$, plus the
following additional structure:
\begin{enumerate}
\item a strong monad $\mathcal{L}$ (lifting) with unit $\eta_X: X \to \mathcal{L}(X)$, multiplication $\mu_X:
\mathcal{L}^2(X) \to \mathcal{L}(X)$ and strength $\mathsf{st}_{X,Y}: X \times \mathcal{L}(Y) \to
\mathcal{L}(X \times Y)$;
\item $\mu$-polynomial endofunctors given by $\muPoly_{\cat{C}}$;
\item for each primitive type $\rho \in \PrimTy$, an object $\sem{\rho}$ of $\cat{C}$;
\item for each primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}:
\sem{\rho_1} \times \ldots \times \sem{\rho_n} \to \sem{\rho}$.
\end{enumerate}

\figref{default-semantics:types} gives the interpretation of types $\Delta \vdash \tau$ with $\length{\Delta}
= n$ as functors $\cat{C}^n \to \cat{C}$, or as objects of $\cat{C}$ in the case $n = 0$.
\figref{default-semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as morphisms
$\sem{\Gamma} \to \sem{\tau}$.

>>>>

To give the semantics for the language defined in \figrefTwo{syntax}{typing}, we fix a bicartesian closed
category $\Sem$ with finite products $(\times, \One)$, finite coproducts $(\coprod, \Zero)$, exponentials
$\internalHom{X}{Y}$, evaluation morphisms $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$, plus the
following additional structure:
\begin{enumerate}
\item a strong monad $\mathcal{L}$ (lifting) with unit $\eta_X: X \to \mathcal{L}(X)$, multiplication $\mu_X:
\mathcal{L}^2(X) \to \mathcal{L}(X)$ and strength $\mathsf{st}_{X,Y}: X \times \mathcal{L}(Y) \to
\mathcal{L}(X \times Y)$;
\item $\mu$-polynomial endofunctors given by $\muPoly_{\Sem}$;
\item for each primitive type $\rho \in \PrimTy$, an object $\sem{\rho}$;
\item for each primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}:
\sem{\rho_1} \times \ldots \times \sem{\rho_n} \to \sem{\rho}$.
\end{enumerate}

\figref{default-semantics:types} gives the interpretation of types $\Delta \vdash \tau$ with $\length{\Delta}
= n$ as functors $\Sem^n \to \Sem$, or as objects of $\Sem$ in the case $n = 0$.
\figref{default-semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as morphisms
$\sem{\Gamma} \to \sem{\tau}$.
